---
title: "Class 7 Git and Data Import"
output: html_document
date: "2026-07-12"
---

# Assignment

Create a repository on github 

Create a project and choose git for the repository 

Version Control --> Git --> copy the repository url 

Create a rmarkdown file, and use the commit tab to save your progress. 

Push your assignment to the repository once completed. 

!!! Please submit your assignment this week by sharing your git repository link with me!!!


## Import the dataset 

Download both datasets from Github 
Import the dataset to your work environment 


```{r}

```

# Change the name of your dataset into snake_case 

```{r}

```


# Explore your data and answer the following questions 

1. How many line of data are there? 

2. List out all the name of the variables within the dataset. Provide me with a short description regarding what each variable. 

- What is the class of the variable?
- Is there any variable missing?
- If this is a categorical variable, please describe what are the categories?
- If this is a numerical variable, tell me what is the mean, what is the median, minimum and maximum?




