---
title: "Class 6 Introduction to R"
output: html_document
date: "2026-07-04"
---

# 1. Introduction to R and R studio 

## Introduction to R studio 

Windows (Console, terminal)
Different r files (Rscript vs. Rmarkdown)
How to run your data 

Insert a code chunk in r markdown 

```{r}

```



# 2. Basic Structure of R Code

## Comments

Anything after `#` is ignored by R.

```{r}
# This is a comment

# Getting the output for the calculation 2+2 

2 + 2
```

---

## Running Code

```{r}
5 + 7

10 * 3

25 / 5
```

---

## Case Sensitivity

R is case sensitive.

```{r}
age <- 25

Age <- 30

age

Age

age <- 15

age
```

Notice these are two different variables.

---

# 3. Naming Variables

## Assignment

The preferred assignment operator is `<-`.

```{r}
age <- 25

height <- 170

name <- "Alice"
```

You may also see:

```{r}
weight = 60

```

Although this works, using `<-` is considered best practice.

---

## Valid Variable Names

```{r}
patient_age <- 45

patientAge <- 45

age2 <- 20
```

snake_case naming format is preferred 
- all lowercase 
- connect with underline

```{r}
patient_age <- 35

blood_pressure <- 120
```


---

## Invalid Variable Names

These examples will produce errors.

```{r}
2age <- 20

patient age <- 20

TRUE <- 1
```


---

# 4. Common Symbols in R

## Assignment

```{r}
x <- 5
```

---

## Arithmetic Operators

| Symbol | Meaning |
|--------|---------|
| + | Addition |
| - | Subtraction |
| * | Multiplication |
| / | Division |
| ^ | Exponent |

Example:

```{r}
5 + 3

10 - 2

4 * 5

20 / 4

2^3
```

---

## Comparison Operators

| Symbol | Meaning |
|--------|---------|
| == | Equal to |
| != | Not equal to |
| > | Greater than |
| < | Less than |
| >= | Greater than or equal |
| <= | Less than or equal |

```{r}
age <- 25

age > 18

age == 25

age != 30
```

**Important**

=     assigns a value

==    compares two values'

---

## Logical Operators

| Symbol | Meaning |
|--------|---------|
| & | AND |
| \| | OR |
| ! | NOT |

```{r}
age <- 25

height <- 170

age > 18 & height > 160

age > 30 | height > 160

!TRUE
```

---

## Parentheses

Used for:

- Functions
- Order of operations

```{r}
(2 + 3) * 4

mean(c(1,2,3,4))
```

---

## Square Brackets

Used to select elements.

```{r}
x <- c(10,20,30,40)

x[1]

x[2]

x[c(1,3)]
```

---

# 5. Basic Data Types

## Numeric

```{r}
age <- 25
```

---

## Character

```{r}
name <- "Alice"
```

Always put character in quotes

---

## Logical

```{r}
passed <- TRUE

notpass <- FALSE
```

---

## Vector

```{r}
scores <- c(85,90,78,92)
```

---

## Check the Data Type

class(variable_of_interest)

```{r}
class(age)

class(name)

class(scores)

class(passed)
```

---

# 6. Useful Functions

```{r}
scores <- c(85,90,78,92)

length(scores)

sum(scores)

mean(scores)

max(scores)

min(scores)
```


---

# 7. Common Errors

## Missing Quotation Mark/ Parenthesis


```{r}
name <- "Alice

```


```{r}
mean(((((scores))))
```

---

## Using = Instead of ==

```{r}
age = 18

age == 18
```

---

# Practice Exercise

Create the following variables.

```{r}
name <- "John" 

age <- 30

height <- 175

weight <- 70

scores <- c(85,90,78,92)
```


Answer these questions:

1. What is the class of `name`?
2. Is age greater than 18?
3. What is the average score?
4. What is the BMI?

```{r}
class_of_name <- class(name)
print(class_of_name)

age>18

mean(scores)

BMI <- weight /((height/100)^2)
print(BMI)
```


```{r}
class(name)
age > 18
scores <- c(88,91,79,95)

mean(scores)
```



# 8. Install R packages

```{r}
install.packages("ggplot2")
```


# 9. Reference the packages that you are going to use 

```{r setup}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readxl)
library(haven)
```


# Data Import

## import a csv file

```{r}
geo_ds = read_csv("C:/Users/Admin/Documents/Vaccine Hesistancy Research/vaccine_nyc/Main_Routine_Vaccine_Geo.csv")
demo_ds = read_csv("C:/Users/Admin/Documents/Vaccine Hesistancy Research/vaccine_nyc/Main_Routine_Vaccine_Demo.csv")
```


## view the dataset

-  head/tail see the first few or last few rows of the data 
-  view() allow you to view the whole dataset *make sure you don't knit that since it can mess up r-markdown

```{r}
head(demo_ds)
tail(demo_ds, 2)
```

```{r}
summary(demo_ds)
```


## janitor clean name 

clean up names into snake_case
```{r}
demo_ds = janitor::clean_names(demo_ds)
head(demo_ds)
```



# Assignment 

## Installation of Git 

Create a github account


https://happygitwithr.com/

Follow the guideline on this website to learn about what is Git and have get Installed on your computer. Read and complete step 1 to 9 


## Install the following packages in R

ggplot2, dplyr, tidyverse, readxl, haven


